Story Cloze Task: UW NLP System

### 2017-02-16 docs/shared_task_paper_camera_ready.tex
Noah, polishing

\newcommand{\ms}[1]{{\color{cyan}\{\textit{#1}\}$_{ms}$}}
\newcommand{\roy}[1]{{\color{orange}\textsc{[#1 --rs]}}}
\newcommand{\royb}[2]{{\color{red}{\sout{#1}}}{\color{green}{#2}}}
\newcommand{\royc}[3]{\royb{#1}{#2}\roy{#3}}
\newcommand{\yc}[1]{{\color{bblue}\{\textit{#1}\}$_{yc}$}}

\newcommand{\tabref}[1]{Table~\ref{#1}}
\newcommand{\com}[1]{}
\definecolor{forestgreen}{rgb}{0.13, 0.55, 0.13}

\title{Story Cloze Task --- UW System}

\author{\hspace{2cm}Roy Schwartz$^{1,2}$ \And \hspace{2.5cm}Maarten Sap$^1$ \And \hspace{3cm}Ioannis Konstas$^1$ \And \\
  $^1$Computer Science \& Engineering, University of Washington, WA 98195, USA \\
  $^2$Allen Institute for Artificial Intelligence, WA 98103, USA \\
    {\tt \{roysch,msap,ikonstas,lzilles,yejin,nasmith\}@cs.washington.edu}  \\
    \And \hspace{-2cm}Leila Zilles$^1$ \And \hspace{-2.5cm}Yejin Choi$^1$ \And \hspace{-2.5cm}Noah A. Smith$^1$ \\
  }

@@

\begin{document}
\maketitle
\begin{abstract} % This is my real quick try at this lol
This paper describes the university of Washington NLP group submission for the Linking Models of Lexical, Sentential and Discourse-level Semantics (LSDSem 2017) shared task---the \textit{Story Cloze Task}.
Our system feeds a linear classifier with a variety of features, including both the output of a neural language model and style features.
We report  $75.2\%$ accuracy on the task. 
A further discussion of our results can be found in \cite{Schwartz:2017}.
\end{abstract}

@@

% Last paragraph should be a quick description of our system.
In this paper, we describe the university of Washington NLP group submission for the shared task.
Our system explores several types of features for the task.
First, we train a neural language model \cite{mikolov2010recurrent} on the ROC story corpus. 
We use the probabilities assigned by the model to each of the endings ({\it right} and {\it wrong}) as classification features.

Second, we attempt to distinguish between {\it right} and {\it wrong} endings using style features, such as sentence length, character n-grams and word n-grams. 
Our intuition is that the {\it right} endings use a different style compared to the {\it wrong} endings.
The features we use were shown useful for style detection in tasks such as age \cite{Schler:2006}, gender \cite{Argamon:2003}, 
and authorship profiling \cite{Stamatatos:2009}.

We feed our features to  a logistic regression classifier, and evaluate our system on the shared task.
Our system obtains $75.2\%$ accuracy on the test set.
Our findings hint that the different writing tasks used to create the story cloze task---writing {\it right} and {\it wrong} endings---impose different writing styles on authors. 
This is further discussed in \cite{Schwartz:2017}.

\begin{table}[!t]
%\small
\begin{tabular}{|p{3.5cm}|p{3.5cm}|} \hline
{\bf Story Prefix} & {\bf Ending} \\ \hline
\multirow{2}{*}{\parbox[b][7.5em][c]{3.5cm}{Kathy went shopping. She found a pair of great shoes. The shoes were \$300. She bought the shoes.}}	&{\color{forestgreen}{She felt buyer's remorse after the purchase.}} \\ \cline{2-2}
& {\color{red}{Kathy hated buying shoes.}} \\ \hline
\end{tabular}
\caption{\label{ROC-example}
Examples of stories from the story cloze task \cite{Mostafazadeh:2016}. 
The left column shows that first four sentences of a story. 
The right column shows two contrastive endings for the story:
 %of the same story: 
 a {\color{forestgreen}{coherent}} ending (upper row) and a {\color{red}{incoherent}} one (bottom row).
}
%\end{center}
\end{table}
\com{
\section{Task summary}
% this section helps clarify some of the nomenclature?

@@

\paragraph{Language Model Features.}
%% take the RLM paragraph from the paper
We experiment with state-of-the-art text comprehension models, specifically an LSTM \cite{hochreiter1997long} recurrent neural network language model (RNNLM; \newcite{mikolov2010recurrent}).
Our RNNLM is used to generate two different probabilities:
$p_\theta(\textrm{ending})$, which is the language model probability of the fifth sentence alone and $p_\theta(\textrm{ending} \mid \textrm{story})$, which is the RNNLM probability of the fifth sentence given the first four sentences.  
We use both of these probabilities as classification features.

@@

\paragraph{Stylistic Features.}
We hypothesize that {\it right} and {\it wrong} endings might be distinguishable using style features.
We adopt features that have been shown useful in the past in tasks such as detection of age \cite{Schler:2006,Rosenthal:2011,nguyen:2011:latech}, gender  \cite{Argamon:2003,Schler:2006,bamman2014gender}, and native language
\cite{Koppel:2005,Tsur:2007,Bergsma:2012}.

We add the following classification features to capture style differences between the two endings. 
These features are computed on the story endings alone ({\it right} or {\it wrong}), and do not consider, neither at train nor at test time, the first four (shared) sentences of each story.
\begin{itemize}
\item\textit{\textbf{Length}.} The number of words in the sentence.
\item\textit{\textbf{Word n-grams.}} We use sequences of 1-5 words. Following \cite{Tsur:2010,Schwartz:2013}, we distinguish between high frequency and low frequency words. 
Specifically, we replace content words, which are often low frequency, with their part-of-speech tags (Nouns, Verbs, Adjectives and Adverbs).
\item\textit{\textbf{Character n-grams.}} Character n-grams are useful features in the detection of author style \cite{Stamatatos:2009} or language identification \cite{lui2011cross}.
We use character 4-grams.
\end{itemize}

@@

We start by tokenizing the text using the nltk tokenizer.\footnote{\url{www.nltk.org/api/nltk.tokenize.html}} 
We then use TensorFlow\footnote{\url{www.tensorflow.org}} to train the RNNLM using a single-layer LSTM of hidden dimension 512.
We use the ROC Stories for training, setting aside 10\% for validation of the language model.\footnote{We train on both the Spring 2016 and the Winter 2017 datasets, a total of roughly 100K stories}
We replace all words occurring less than 3 times by a special out-of-vocabulary character, yielding a vocabulary size of 21,582.
Only during training, we apply a dropout rate of 60\% while running the LSTM over all 5 sentences of the stories. 
Using Adam optimizer \cite{kingma2014adam} and a learning rate of
$\eta=.001$, we train to minimize cross-entropy.

For computing the style features, the story cloze task doesn't have a training corpus for the {\it right} and {\it wrong} endings.
Therefore, we use the development set as our training set, holding out 10\% for development
(3,366 training endings, 374 for development). We keep the story cloze test set as is (3,742 endings).
We add a \textsc{start} symbol at the beginning
of each sentence.\footnote{Virtually all sentences end with a period
  or an exclamation mark, so we do not add a \textsc{stop} symbol.} 
 We keep n-gram (character or word) features that occur at least five times in the training set.
All stylistic feature values are normalized to [0-1].
For the POS features, we tag all endings with the Spacy POS tagger.\footnote{\url{spacy.io/}}
The total number of features used by our system is 7,651.

@@

\begin{table}%[!t]
\begin{center}
%\small
\begin{tabular}{|l|r|} \hline
{\bf Model} & {\bf Acc.} \\ \hline
{DSSM} \cite{Mostafazadeh:2016} & 0.585 \\ 
{LexVec} \cite{Salle:2016} & 0.599 \\ \hline\hline
%{Niko (shared task)}	& 0.700\\ 
%{tbmihaylov (shared task)} & 0.711\\ \hline\hline
{RNNLM Features}		& 0.677 \\ 
{Stylistic features} & {0.724} \\ 
{\bf Combined (Style + RNNLM)} & {\bf 0.752} \\ \hline\hline
Human judgment & 1.000 \\ \hline
\end{tabular}
\end{center}
\caption{\label{cloze_results}
Results on the test set of the  story cloze task. 
The first block are published results, the second block are our results.
LexVec results are taken from \cite{Speer:2016}.
Human judgement scores are taken from \cite{Mostafazadeh:2016}. 
}
\end{table}

@@

In an attempt to provide explanation to the strong performance of the stylistic feature, 
we hypothesize that the different writing tasks---writing a {\it right} and a {\it wrong} ending---impose a different style on the authors, which is expressed in the different style adopted in each of the cases. 
The reader is referred to \cite{Schwartz:2017} for more details and discussion.

\section{Conclusion}
This paper described the university of Washington NLP group submission to the LSDSem 2017 Shared Task. 
Our system leveraged both neural language model features and  stylistic features, achieving $75.2\%$ accuracy on the classification task.

>>>>

\newcommand{\ms}[1]{{\color{cyan}\{\textit{#1}\}$_{ms}$}}
\newcommand{\roy}[1]{{\color{orange}\textsc{[#1 --rs]}}}
\newcommand{\royb}[2]{{\color{red}{\sout{#1}}}{\color{green}{#2}}}
\newcommand{\royc}[3]{\royb{#1}{#2}\roy{#3}}
\newcommand{\yc}[1]{{\color{bblue}\{\textit{#1}\}$_{yc}$}}
\newcommand{\nascomment}[1]{\textcolor{blue}{[\textsc{#1 --nas}]}}
\newcommand{\tabref}[1]{Table~\ref{#1}}
\newcommand{\com}[1]{}
\definecolor{forestgreen}{rgb}{0.13, 0.55, 0.13}

\title{Story Cloze Task: UW NLP System}

\author{\hspace{2cm}Roy Schwartz$^{1,2}$ \And \hspace{2.5cm}Maarten Sap$^1$ \And \hspace{3cm}Ioannis Konstas$^1$ \And \\
  $^1$Computer Science \& Engineering, University of Washington,
  Seattle, WA 98195, USA \\
  $^2$Allen Institute for Artificial Intelligence, Seattle, WA 98103, USA \\
    {\tt \{roysch,msap,ikonstas,lzilles,yejin,nasmith\}@cs.washington.edu}  \\
    \And \hspace{-2cm}Leila Zilles$^1$ \And \hspace{-2.5cm}Yejin Choi$^1$ \And \hspace{-2.5cm}Noah A.~Smith$^1$ \\
  }

@@

\begin{document}
\maketitle
\begin{abstract} % This is my real quick try at this lol
This paper describes University of Washington NLP's submission for the Linking Models of Lexical, Sentential and Discourse-level Semantics (LSDSem 2017) shared task---the \textit{Story Cloze Task}.
Our system is a linear classifier with a variety of features, including both the scores of a neural language model and style features.
We report  $75.2\%$ accuracy on the task. 
A further discussion of our results can be found in \newcite{Schwartz:2017}.
\end{abstract}

@@

% Last paragraph should be a quick description of our system.
In this paper, we describe University of Washington NLP's submission for the shared task.
Our system explores several types of features for the task.
First, we train a neural language model \cite{mikolov2010recurrent} on the ROC story corpus. 
We use the probabilities assigned by the model to each of the endings ({\it right} and {\it wrong}) as classification features.

Second, we attempt to distinguish between {\it right} and {\it wrong} endings using style features, such as sentence length, character $n$-grams and word $n$-grams. 
Our intuition is that the {\it right} endings use a different style compared to the {\it wrong} endings.
The features we use were shown useful for style detection in tasks such as age \cite{Schler:2006}, gender \cite{Argamon:2003}, 
and authorship profiling \cite{Stamatatos:2009}.

We feed our features to  a logistic regression classifier, and evaluate our system on the shared task.
Our system obtains $75.2\%$ accuracy on the test set.
Our findings hint that the different writing tasks used to create the story cloze task---writing {\it right} and {\it wrong} endings---impose different writing styles on authors. 
This is further discussed in \newcite{Schwartz:2017}.

\begin{table}[!t]
%\small
\begin{tabular}{|p{3.5cm}|p{3.5cm}|} \hline
{\bf Story Prefix} & {\bf Ending} \\ \hline
\multirow{2}{*}{\parbox[b][7.5em][c]{3.5cm}{Kathy went shopping. She found a pair of great shoes. The shoes were \$300. She bought the shoes.}}	&{\color{blue}{She felt buyer's remorse after the purchase.}} \\ \cline{2-2}
& {\color{red}{Kathy hated buying shoes.}} \\ \hline
\end{tabular}
\caption{\label{ROC-example}
Examples of stories from the story cloze task \cite{Mostafazadeh:2016}. 
The left column shows that first four sentences of a story. 
The right column shows two contrastive endings for the story:
 %of the same story: 
 a {\color{blue}{coherent}} ending (upper row) and a {\color{red}{incoherent}} one (bottom row).
}
%\end{center}
\end{table}
\com{
\section{Task summary}
% this section helps clarify some of the nomenclature?

@@

\paragraph{Language model features.}
%% take the RLM paragraph from the paper
We experiment with state-of-the-art text comprehension models,
specifically an LSTM \cite{hochreiter1997long} recurrent neural
network language model (RNNLM; \nocite{mikolov2010recurrent}Mikolov et
al., 2010).
Our RNNLM is used to generate two different probabilities:
$p_\theta(\textrm{ending})$, which is the language model probability of the fifth sentence alone and $p_\theta(\textrm{ending} \mid \textrm{story})$, which is the RNNLM probability of the fifth sentence given the first four sentences.  
We use both of these probabilities as classification features.

@@

\paragraph{Stylistic features.}
We hypothesize that {\it right} and {\it wrong} endings might be distinguishable using style features.
We adopt features that have been shown useful in the past in tasks such as detection of age \cite{Schler:2006,Rosenthal:2011,nguyen:2011:latech}, gender  \cite{Argamon:2003,Schler:2006,bamman2014gender}, and native language
\cite{Koppel:2005,Tsur:2007,Bergsma:2012}.

We add the following classification features to capture style differences between the two endings. 
These features are computed on the story endings alone ({\it right} or {\it wrong}), and do not consider, either at train nor at test time, the first four (shared) sentences of each story.
\begin{itemize}
\item\textit{\textbf{Length}.} The number of words in the sentence.
\item\textit{\textbf{Word $n$-grams.}} We use sequences of 1--5
  words. Following \newcite{Tsur:2010} and \newcite{Schwartz:2013}, we distinguish between high frequency and low frequency words. 
Specifically, we replace content words, which are often low frequency, with their part-of-speech tags (Nouns, Verbs, Adjectives, and Adverbs).
\item\textit{\textbf{Character $n$-grams.}} Character $n$-grams are useful features in the detection of author style \cite{Stamatatos:2009} or language identification \cite{lui2011cross}.
We use character 4-grams.
\end{itemize}

@@

We start by tokenizing the text using the nltk tokenizer.\footnote{\url{www.nltk.org/api/nltk.tokenize.html}} 
We then use TensorFlow\footnote{\url{www.tensorflow.org}} to train the RNNLM using a single-layer LSTM of hidden dimension 512.
We use the ROC Stories for training, setting aside 10\% for validation of the language model.\footnote{We train on both the Spring 2016 and the Winter 2017 datasets, a total of roughly 100K stories.}
We replace all words occurring less than 3 times by a special out-of-vocabulary character, yielding a vocabulary size of 21,582.
Only during training, we apply a dropout rate of 60\% while running the LSTM over all 5 sentences of the stories. 
Using Adam optimizer \cite{kingma2014adam} and a learning rate of
$\eta=.001$, we train to minimize cross-entropy.

For computing the style features, the story cloze task doesn't have a training corpus for the {\it right} and {\it wrong} endings.
Therefore, we use the development set as our training set, holding out 10\% for development
(3,366 training endings, 374 for development). We keep the story cloze test set as is (3,742 endings).
We add a \textsc{start} symbol at the beginning
of each sentence.\footnote{Virtually all sentences end with a period
  or an exclamation mark, so we do not add a \textsc{stop} symbol.} 
 We keep $n$-gram (character or word) features that occur at least five times in the training set.
All stylistic feature values are normalized to [0, 1] \nascomment{the
  range [0, 1] or the two values \{0, 1\}?}
For the part-of-speech features, we tag all endings with the Spacy POS tagger.\footnote{\url{spacy.io/}}
The total number of features used by our system is 7,651.

@@

\begin{table}%[!t]
\begin{center}
%\small
\begin{tabular}{|l|r|} \hline
{\bf Model} & {\bf Acc.} \\ \hline
{DSSM} \cite{Mostafazadeh:2016} & 0.585 \\ 
{LexVec} \cite{Salle:2016} & 0.599 \\ \hline\hline
%{Niko (shared task)}	& 0.700\\ 
%{tbmihaylov (shared task)} & 0.711\\ \hline\hline
{RNNLM features}		& 0.677 \\ 
{Stylistic features} & {0.724} \\ 
{\bf Combined (Style + RNNLM)} & {\bf 0.752} \\ \hline\hline
Human judgment & 1.000 \\ \hline
\end{tabular}
\end{center}
\caption{\label{cloze_results}
Results on the test set of the  story cloze task. 
The first block are published results, the second block are our results.
LexVec results are taken from \cite{Speer:2016}.
Human judgement scores are taken from \cite{Mostafazadeh:2016}. 
}
\end{table}

@@

In an attempt to provide explanation to the strong performance of the stylistic feature, 
we hypothesize that the different writing tasks---writing a {\it right} and a {\it wrong} ending---impose a different style on the authors, which is expressed in the different style adopted in each of the cases. 
The reader is referred to \newcite{Schwartz:2017} for more details and discussion.

\section{Conclusion}
This paper described University of Washington NLP's submission to the LSDSem 2017 Shared Task. 
Our system leveraged both neural language model features and  stylistic features, achieving $75.2\%$ accuracy on the classification task.

\section*{Acknowledgments}
The authors thank the shared task organizers and anonymous reviewers
for feedback.  \nascomment{anyone else} This research was supported in
part by the Allen Institute for Artificial Intelligence and in part by
DARPA under the Communicating with Computers program.
